---
title: "Untitled"
output: html_document
---

# Question 5 {-}
Suppose this data represents the population of newborns in one hospital and you take a simple random sample of 100 babies from the population.

a) What is the probability that your sample contains more than 25 babies has abnormal health status (1 = health, 2 = suspect, 3 = Pathological)? 

```{r}
## ---- Question-5a ------------------------------------------------------------

df_fh <- readr::read_csv(here::here("fetal_health.csv"))
library(dplyr)
library(tidyr)
library(mosaic)

df_fh %>%
  select(fetal_health)%>%
  filter(fetal_health != 1)%>%
  nrow()

#there are 471 abnormal in this population

#the probability of having abnormal is 471/2126 = 0.22

1 - mosaic::xpbinom(q = 25, size = 100, prob = 0.22)

```
The probability of having more than 30 abnormal is 0.197. 

b) Turns out that your sample actually contains 20 babies with abnormal health status. What is the 95% confidence interval of this proportion? Can you use a normal approximation for this sample? Why or why not?
```{r}
## ---- Question-5b ------------------------------------------------------------

mosaic::binom.test(x = 200, n = 1000, ci.method = "Clopper-Pearson")
mosaic::binom.test(x = 200, n = 1000, ci.method = "Wald")

mosaic::binom.test(x = 20, n = 100, ci.method = "Clopper-Pearson")
mosaic::binom.test(x = 20, n = 100, ci.method = "Wald")
```
The 95% CI using Clopper-Pearson method: [0.1266556,0.2918427]
The 95% CI using exact method(ie: normal approximation): [0.1216014,0.2783986]
The two methods give similar 95% CIs. Normal approximation can be used here since the sample size is large enough to generate a binomial distribution approximating the normal distribution and for CLT to kick in. 

c) Another sample taken have the same proportion of event but the sample size is now only 10 and the count of abnormal is 2. Calculate the 95% CI using this sample and compare it with the one you have in b). Describe their difference and the reason why. 
```{r}
## ---- Question-5c ------------------------------------------------------------

mosaic::binom.test(x = 2, n = 10, ci.method = "Clopper-Pearson")
```
The 95% CI for this sample is: [0.02521073, 0.55609546].
95% CI in c) with sample size = 10 is wider than the 95% CI in b) with sample size = 100. 
The sample size is different so the standard error is different. The 95% CI is calculated using the formula: 
$$
\bar y - 1.96(\frac {\sigma} {\sqrt n}) , \bar y + 1.96(\frac {\sigma} {\sqrt n}) 
$$
If the sample size n is larger, the standard error(sigma/sqrt n) is smaller. The value 1.96*standard error is also smaller, and results in a narrower confident interval.


\newpage


\vspace{1.5in}

```{r, echo=TRUE}
## --------------------------- End -------------------------------
```



\newpage


# Code {-}

```{r all-code, ref.label=knitr::all_labels(), echo = TRUE, eval = FALSE}
# Attach this code to Question 7
```

