---
title: "Q4"
author: "Xinbei Wan"
date: "14/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 4. p-value, power {-}

For the purpose of this question only, we treat the 2126 individuals as **the entire target population of newborns**. 

## 1) Calculate the mean and standard deviation of the baseline fetal heart rate.

```{r}
fetal_health <- readr::read_csv(here::here("fetal_health.csv"))
mean_hr <- mean(fetal_health$`baseline value`)
mean_hr
sd_hr <- sqrt(var(fetal_health$`baseline value`)*(length(fetal_health$`baseline value`)-1)/length(fetal_health$`baseline value`))
sd_hr
```

## 2) It is claimed that the baseline fetal heart rates of fetuses with "suspect" health status are above average. Take a simple random sample of 10 fetuses with "suspect" health status, and measure their heart rate to obtain a sample mean of 141.68. Heart rates are scaled to be normally distributed. Does the sample provide evidence to reject null hypothesis? State your null and alternative hypothesis.

```{r}
pnorm(q = 141.68, mean = 133.30, sd = 9.84/sqrt(10), lower.tail = FALSE)
```

$H_0: \mu = 133.30$, $H_A: \mu > 133.30$

The p-value of one sided test is 0.0035. This sample provides evidence against the null hypothesis. The p-value tells us the probability of observing the sample size mean of 141.68 under the null hypothesis distribution is very unlikely.

## 3) What power do you have to detect the baseline fetal heart rates of fetuses with "suspect" health status are at least 8.38 heart beats higher than average, using a one-sided test and sample size 10 and a 0.05 level test?

$H_0: \mu = 133.30$, $H_A: \mu > 141.30$
```{r}
# cutoff to reject the null
cutoff <- qnorm(p = 0.95, mean = 133.30, sd = 9.84/sqrt(10))
cutoff
# probability of observing this cutoff or greater under the alternative
pnorm(q = cutoff, mean = 141.68, sd = 9.84/sqrt(10), lower.tail = FALSE)
```

## 4) A sample size of 10 fetuses with "suspect" health status will have at least 85% power to detect a difference of 8.38 heart beats. Use a simulation based approach to reproduce the sample size calculation for the baseline fetal heart rates of fetuses with "suspect" health status and average.

```{r}
set.seed(490)

power_distribution <- replicate(n = 1000, expr={
  sample.size <- 10
  
  suspect <- rnorm(sample.size, mean = 141.68, sd = 9.83)
  SEM <- sd(suspect)/sqrt(sample.size)
  
  pnorm(q = mean(suspect), mean = 133.30, sd =SEM, lower.tail = FALSE) < 0.05

} )

prop.table(table(power_distribution))
```

The percentage of samples that results in a p-value less than 0.05 is 85.4%, which shows the study is powered at 85% to detect the difference.


